---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{R loadAndProcessData}
library(ggplot2)
library(plyr)
library(reshape2)
library(xtable)
setwd("C:/DataScience/Reproducible Research/W2 Project/RepData_PeerAssessment1/");

activityFile <- "activity.zip"
activityData <- read.csv(unz(activityFile, "activity.csv"), na.strings = "NA")

#fixed date type, only realized this error later.
activityData$date <- as.Date(as.character(activityData$date))
```


## What is mean total number of steps taken per day?

### histogram of the steps 
```{r plot,results="asis"}
steps <- tapply(activityData$steps, activityData$date, FUN = sum, na.rm = TRUE)
qplot(steps, binwidth = 1000, xlab = "total number of steps taken each day") 
```

### Mean Steps per date
```{r results="asis"}
stepsperDay <-aggregate(steps ~ date, data=activityData, sum, na.action = na.omit)
colnames(stepsperDay) <-c("date","mean steps")

mean(stepsperDay$`mean steps`)

```

## What is the average daily activity pattern?
### Average Activity per day
```{r results="asis"}
ave<- aggregate(x = list(steps = activityData$steps), by = list(interval = activityData$interval), 
    FUN = mean, na.rm = TRUE)
ggplot(data = ave, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute intervals") + 
    ylab("average number of steps taken")

```

### The average daily Activity is 206.1698
```{r}
ave[which.max(ave$steps ),]
```

## Imputing missing values

###There are quite a few missing=NA values. We need to replace those missing values with the mean steps in the associate interval. I am using the mean imputing method reference here [https://en.wikipedia.org/wiki/Imputation_(statistics)]

```{r }
missingSteps<-is.na(activityData$steps)

table(missingSteps)
```


```{r imputingUsingMean}

fill_value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (ave[ave$interval==interval, "steps"])
    return(filled)
}

activityFilled <- activityData
activityFilled$steps <- mapply(fill_value, activityFilled$steps, activityFilled$interval)

```

###Lets plot this filled dataset

```{r}

totalNumSteps <- tapply(activityFilled$steps, activityFilled$date,FUN=sum)
qplot(totalNumSteps, binwidth = 1000, xlab = "Total number of steps per day")
```

## Are there differences in activity patterns between weekdays and weekends?

###Seperate the data between weekend and weekday
```{r }
dayofWeek <- ifelse(weekdays(activityFilled$date)=="Saturday" | weekdays(activityFilled$date)=="Sunday","weekend","weekday")
activityFilled$day <- as.factor(dayofWeek)

weekdayActivities<- subset(activityFilled,day == "weekday")
wdActivity <- ddply(weekdayActivities, c("interval","day"), summarise, avgSteps=mean(steps))

weekendActivities<- subset(activityFilled,day == "weekend")

weActivity <- ddply(weekendActivities, c("interval","day"), summarise, avgSteps=mean(steps))

```
